Extending Camellia for Distributed Stiffness Matrix Determination in DPG

### 2011-12-09 docs/HPC_report/Extending_Camellia.tex
added figure, more explanation to report
git-svn-id: https://svn.ices.utexas.edu/repos/camellia@105 36274a68-8941-4c93-80ef-affaed5139da

\documentclass{article}
\newcommand{\oneFig}{.65}
\newcommand{\twoFig}{.42}
\newcommand{\fourFig}{.42}

@@

We tested the scalability of our code on the Lonestar machine at the Texas Advanced Computing Center (TACC), verifying whether or not we observe strong/weak scaling for various parts of the program.  We solve the convection-diffusion equation on a square mesh, with boundary conditions such that the solution develops a boundary layer (sharp gradient) near the North and East edges of the square.  We set up our computational experiments by first refining all cells that share an edge with the North and East sides, and repeating this process four times.   We then construct three additional meshes, by uniformly refining every element in the mesh (quadrupling the number of elements each time).  For each of these meshes, we solve on Longhorn, using 1, 4, 16, and 64 nodes, tracking the timing of various portions of the code.  From this data, we extract strong and weak scaling statistics.

@@

\begin{figure}
\centering
\subfigure[Global stiffness assembly]{
\includegraphics[scale = \twoFig]{figs/scalingFigs/weakScalingAssembly.pdf}
}
\subfigure[Local stiffness computation]{
\includegraphics[scale = \twoFig]{figs/scalingFigs/weakScalingLocal.pdf}
}
\end{figure}

>>>>

\documentclass{article}
\newcommand{\oneFig}{.5}
\newcommand{\twoFig}{.42}
\newcommand{\fourFig}{.42}

@@

We tested the scalability of our code on the Lonestar machine at the Texas Advanced Computing Center (TACC), verifying whether or not we observe strong/weak scaling for various parts of the program.  We solve the convection-diffusion equation on a square mesh, with boundary conditions such that the solution develops a boundary layer (sharp gradient) near the North and East edges of the square.

\begin{figure}
\centering
\includegraphics[scale=\oneFig]{figs/Solution3232nomesh.png}
\caption{Computed convection-diffusion solution for $\epsilon=1e-2$ using a refined mesh.  $L^2$ error is $O(1e-4)$.}
\end{figure}

We set up our computational experiments by first refining all cells that share an edge with the North and East sides, and repeating this process four times.   We then construct three additional meshes, by uniformly refining every element in the mesh (quadrupling the number of elements each time).  For each of these meshes, we solve on Longhorn, using 1, 4, 16, and 64 nodes, tracking the timing of various portions of the code.  From this data, we extract strong and weak scaling statistics.

We experimented also with both a localized and non-localized partitioning of each mesh in order to analyze the communication costs in global stiffness matrix assembly.  For these runs, we created localized partitions of the mesh using a HSFC partitioner.  For a non-localized partitioner, we chose to use the Zoltan cyclic partitioner, which loops through the list of active elements, sending the first element to the first processor, the next to the next processor, and so on.  After reaching the last processor, the cyclic partitioner cycles back to the first processor and repeats this process \footnote{Due to the nature of our refinements in this test, a blocked algorithm (equally partitioning the ordered list of active elements by element number) still maintained fairly good locality.}.  This partitioning is, in some sense, a ``worst case" scenario that maximizes communication between processors.

@@

\begin{figure}
\centering
\subfigure[Global stiffness assembly]{
\includegraphics[scale = \twoFig]{figs/scalingFigs/weakScalingAssembly.pdf}
}
\subfigure[Local stiffness computation]{
\includegraphics[scale = \twoFig]{figs/scalingFigs/weakScalingLocal.pdf}
}
\caption{Weak scaling results for computation of local stiffness matrices and global stiffness matrix assembly.  For each run, both the number of MPI nodes and number of elements are increased by a factor of four.}
\end{figure}

Both localized and non-localized local stiffness computation display weak scalability by maintaining a constant runtime as both the problem size and number of processors are increased.  Unlike the localized case, non-localized assembly of the global stiffness matrix does not scale weakly. It's worth noting, however, that though communication costs grow much faster with a poor partitioning of the mesh, the maximum amount of time spent on assembly is still only a fraction of the time spent on computation of local stiffness matrices.
